Synthesising Recursive Functions for First-Order Model Counting

### 2023-03-02 doc/paper/supplementary_material.tex
Clean up the Preliminaries

\title{Synthesising Recursive Functions for First-Order Model Counting}

% TODO (at the very end)
% put the appendix in a separate file
% grammarly the whole thing
% read the formatting instructions and their LaTeX source (email font?)
% make sure there are no overflow errors
% clean up unnecessary things from the header (e.g., packages and definitions)
%  make sure the result table (as well as other floating objects) are on the right pages

\author{%
Paulius Dilkas$^1$\and
Vaishak Belle$^2$\\
\affiliations
$^1$National University of Singapore, Singapore, Singapore\\
$^2$University of Edinburgh, Edinburgh, UK\\
\emails
paulius.dilkas@nus.edu.sg,
vbelle@ed.ac.uk
}

\crefalias{diagram}{equation}
\crefname{diagram}{Diagram}{Diagrams}
\Crefname{diagram}{Diagram}{Diagrams}
\creflabelformat{diagram}{#2\textup{(#1)}#3}

@@

@@

In this appendix, we list the exact function definitions produced by
\textsc{Crane} for all of the problem instances in Section~6 both before and
after algebraic simplification (excluding multiplications by one). The
correctness of all of them has been checked by identifying suitable base cases
and verifying the numerical answers across a range of domain sizes.

>>>>

\crefalias{diagram}{equation}
\crefname{diagram}{diagram}{diagrams}
\Crefname{diagram}{Diagram}{Diagrams}
\creflabelformat{diagram}{#2\textup{(#1)}#3}

@@

TODO: finish arranging the definitions

We begin with a few definitions that formalise the notion of a model.

For any set $S$, let $2^{S}$ denote its power set and
$S^{\ast} \coloneqq \bigcup_{i=0}^{\infty} S^{i}$ the set of tuples of any
finite length of elements of $S$. For example,
$\{\,a, b\,\}^{\ast} = \{\, (), (a), (b), (a, a), \dots \,\}$.

Let $\Preds$ be the function that maps any clause or formula to the set of
predicates used within.

For any $n \in \mathbb{N}_{0}$, let $[n] \coloneqq \{\, 1, \dots, n \,\}$, e.g.,
$[0] = \emptyset$, and $[2] = \{\, 1, 2 \,\}$.

Each formula $\phi$ has a map
$\pi_{\phi}\colon \Preds(\phi) \to {\Doms(\phi)}^{\ast}$ s.t., for each
predicate $\texttt{p}/n \in \Preds(\phi)$, we have that
$\pi_{\phi}(\texttt{p}) \in {\Doms(\phi)}^{n}$.

\begin{definition}\label{def:model}
  Let $\phi$ be a formula and $\sigma$ a domain size function. A \emph{model} of
  $(\phi, \sigma)$ is a map
  $\mathfrak{M}\colon \Preds(\phi) \to 2^{\mathbb{N}_{0}^{\ast}}$ s.t.\ the
  following two conditions are satisfied.
  \begin{enumerate}
    \item Let $\texttt{p}/n \in \Preds(\phi)$ be a predicate and let
          $\pi_{\phi}(\texttt{p}) = {(d_{i})}_{i=1}^{n}$ for some domains
          $d_{i} \in \Doms(\phi)$. Then
          \begin{equation}\label{eq:subset}
            \mathfrak{M}(\texttt{p}) \subseteq \prod_{i=1}^{n}[\sigma(d_{i})].\footnote{For simplicity, \cref{def:model} ignores constants. To include constants, one would replace $[\sigma(d_{i})]$ with a set that contains all constants associated with domain $d_{i}$, extended with enough new elements to make its cardinality $\sigma(d_{i})$.}
          \end{equation}
    \item As a collection of relations, $\mathfrak{M}$ satisfies $\phi$.
    \begin{itemize}
      \item TODO: introduce the substitution for formulas, moving the definition from elsewhere
      \item define inductively as...
      \item (not exhaustive)
      \item $\mathfrak{M} \models \phi \land \psi$ iff $\mathfrak{M} \models \phi$ and $\mathfrak{M} \models \psi$
      \item $\mathfrak{M} \models \phi \lor \psi$ iff $\mathfrak{M} \models \phi$ or $\mathfrak{M} \models \psi$
      \item $\mathfrak{M} \models \neg\phi$ iff $\mathfrak{M} \not\models \phi$
      \item $\mathfrak{M} \models (\forall X \in \Delta\text{. }\phi)$ iff $\mathfrak{M} \models \phi[x/\{\, X \,\}]$ for all $x \in \Delta$
      \item One can then add inequality constraints to the semantics as follows
      \item $\mathfrak{M} \models (\forall X \in \Delta\text{. } X \ne c \Rightarrow \phi)$ iff $\mathfrak{M} \models \phi[x/\{\, X \,\}]$ for all $x \in \Delta \setminus \{\, c \,\}$
      \item $\mathfrak{M} \models (\forall X,Y \in \Delta\text{. } X \ne Y \Rightarrow \phi)$ iff $\mathfrak{M} \models \phi[x/\{\, X \,\}][y/\{\, Y \,\}]$ for all $x,y \in \Delta$ s.t. $x \ne y$
    \end{itemize}
  \end{enumerate}
\end{definition}

@@

Here we list the exact function definitions produced by \textsc{Crane} for all
of the problem instances in Section~6 both before and after algebraic
simplification (excluding multiplications by one). The correctness of all of
them has been checked by identifying suitable base cases and verifying the
numerical answers across a range of domain sizes.
